---
title: "classification_test6"
author: "Natalie"
date: "2024-07-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Load packages, message=F}
library(mlr3)
library(mlr3learners)
library(mlr3spatial)
library(terra)
library(sf)
library(ggplot2)
library(tidyterra)
library(here)
library(mlr3cluster)
library(OpenImageR)
library(EBImage)
```
Import datasets
```{r}
site6_rast <- rast(here("Data", "site6_cropped", "site6_smallER.tif"))
names(site6_rast) <- c("Band_1", "Band_2", "Band_3")
site6_rast

#The geopackage contains 200 random points per polygon and spectral information (rgb values) from the raster. 
site6_vect_rgb <- read_sf(here("Data", "site6_cropped", "site6_cropped_pts_rgb.gpkg"), stringsAsFactors = TRUE)
keep <- c("className", "geom")
site6_vect <- site6_vect_rgb[, keep]
site6_vect
```
# Calculate edges of raster and add it as a layer

Convert the site6 raster to an array
```{r}
site6_array <- as.array(site6_rast)
```

Calculate the number of edges per pixel
```{r}
#convert the array to grayscale
site6_gray <- rgb_2gray(site6_array)
#remove all NA values
site6_gray[is.na(site6_gray)] <- mean(site6_gray, na.rm = TRUE)
#calculate edges
edges <- edge_detection(site6_gray, method = "Sobel")
```

Convert "edges" to a raster layer and ensure it has the same crs and extent
```{r}
edges_rast <- rast(edges)
ext(edges_rast) <- ext(site6_rast)
crs(edges_rast) <- crs(site6_rast)
names(edges_rast) <- "edges"
plot(edges_rast)
```
Add the edges layer to the existing site6 raster
```{r}
site6_combined <- c(site6_rast, edges_rast)
site6_combined
```
#Prepare data for model training

Sample rgb and edge values at the training points of the geopackage
```{r}
#Convert geopackage to spatvector and extract values at points
site6_pts <- vect(site6_vect)
sampled_vals <- extract(site6_combined, site6_pts)
site6_pts <- cbind(site6_pts, sampled_vals)
#convert spatVector into dataframe
site6_pts_df <- as.data.frame(site6_pts)
coords <- geom(site6_pts, df=T,xnm="x", ynm="y")
site6_pts_df <- cbind(site6_pts_df, coords)
site6_pts_df
```
Remove unneeded coumns
```{r}
site6_pts_df <- select(site6_pts_df, -part, -hole, -geom, -ID)
site6_pts_df
```

#Run mlr3 model
Create spatial mlr3 task
```{r}
task <- TaskClassifST$new(id = "classification_task",
                           backend = site6_pts_df,
                           target = "className",
                           coordinate_names = c("x", "y"))
task
```
Partition the task into 70% train and 30% test sets
```{r}
set.seed(100)
splits <- partition(task, ratio = 0.7)
```

Create and train random forest learner
```{r}
learner <- lrn("classif.ranger")
train <- learner$train(task, splits$train)
```

Test learner and obtain overall accuracy
```{r}
test <- learner$predict(task, splits$test)
test$score(msr("classif.acc"))
```
Model accuracy is not too different from if we only use rgb. 

Predict an entire raster using the learner and plot the results
```{r}
classified <- predict_spatial(site6_combined, learner)

ggplot() +
  geom_spatraster(data = classified) +
  scale_fill_viridis_d(name = "Class", labels = c("Antelope Brush", "Ground", "Vegetation")) +
  theme_minimal()
```
Calculate area of each class
```{r}
activeCat(classified) <- "categories"
classified
f <- freq(classified)
f$area <- f$count * (0.013 * 0.013) #in m2
f$percent <- round(100 * f$area / sum(f$area), 1)
f
```

