---
title: "classification_test3"
author: "Natalie"
date: "2024-07-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load packages, message=F}
library(terra)
library(sf)
library(dplyr)
library(ggplot2)
library(EBImage)
library(mlr3)
library(mlr3learners)
library(here)
library(OpenImageR)
library(exactextractr)
library(mlr3spatial)
```

# Preprocessing
Import datasets
```{r}
#import site6 raster
site6 <- rast(here("Data", "TestClassification", "site6_smallER.tif"))
names(site6) <- c("red", "green", "blue") 
#import training polygons
site6_poly <- st_read(here("Data", "TestClassification", "site6_poly_smaller_reprojected.shp"))
#import site boundary
site6_boundary <- st_read(here("Data", "TestClassification", "site6_cropBoundary_smaller.shp"))
```

Extract pixel values from raster to create a list and merge elements together into a dataframe
```{r}
#extract rgb pixel values within training polygons and merge them into a dataframe
rgb_values <- exactextractr::exact_extract(site6, site6_poly, include_cols = "className")
rgb_values_tibble <- as_tibble(rgb_values[[1]])

for (i in 3:length(site6_poly)) {
  temp <- as_tibble(site6_poly[[i]])
  rgb_values_tibble <- bind_rows(rgb_values, temp)
}
#only include pixels that are at least 95% within a polygon
rgb_values_tibble_final <- rgb_values_tibble %>%
  filter(coverage_fraction >= 0.95)
#remove all NA values from the tibble
rgb_values_tibble_final_nonas <- rgb_values_tibble_final %>%
  tidyr::drop_na()
#remove the "geometry" and "coverage_fraction" column
rgb_values_tibble_final_nonas <- select(rgb_values_tibble_final_nonas, -geometry)
rgb_values_tibble_final_nonas <- select(rgb_values_tibble_final_nonas, -coverage_fraction)
#display dataframe
print(rgb_values_tibble_final_nonas)
#rename dataframe 
data <- rgb_values_tibble_final_nonas
```

Convert the className column to a factor so it can be used in the mlr3 task. 
```{r}
data$className <- as.factor(data$className)
```

# mlr3 
Create task
```{r}
task <- as_task_classif(data, target = "className")
task
```
Partition task into 70% training and 30% testing
```{r}
set.seed(100)
splits <- partition(task, ratio = 0.7)
```

Create random forest learner
```{r}
learner <- mlr3::lrn("classif.ranger", predict_type = "prob")
```
Train random forest learner
```{r}
training <- learner$train(task, splits$train)
```

Validate random forest learner and obtain overall accuracy
```{r}
testing <- learner$predict(task, splits$test)
#obtain overall accuracy of model
testing$score(msr("classif.acc"))
```

# Apply model to new dataset

Extract rgb values of all cells, including their coordinates, and create a dataframe
```{r}
rgb_values_wholesite <- extract(site6, site6_boundary, xy=T)
print(rgb_values_wholesite)

#convert to dataframe
rgb_values_wholesite_df <- as.data.frame(rgb_values_wholesite)
rgb_values_wholesite_df
```
Remove NA values
```{r}
rgb_values_wholesite_nonas <- rgb_values_wholesite_df %>%
  tidyr::drop_na()
print(rgb_values_wholesite_nonas)
```

Add "className" column and convert it to a factor
```{r}
rgb_values_wholesite_nonas$className <- NA
rgb_values_wholesite_nonas$className <- as.factor(rgb_values_wholesite_nonas$className)
print(rgb_values_wholesite_nonas)
```
Apply the model to a new dataframe
```{r}
predicted_classes <- learner$predict_newdata(newdata = rgb_values_wholesite_nonas)
print(predicted_classes)
```

