---
title: "classification_test2"
author: "Natalie"
date: "2024-07-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=F}
library(terra)
library(sf)
library(dplyr)
library(ggplot2)
library(EBImage)
library(mlr3)
library(mlr3learners)
library(here)
library(OpenImageR)
library(exactextractr)
library(mlr3spatial)
```

```{r}
site6 <- rast(here("Data", "TestClassification", "site6_smallER.tif"))
names(site6) <- c("red", "green", "blue")

site6_poly <- st_read(here("Data", "TestClassification", "site6_poly_smaller_reprojected.shp"))
```

Extract pixel values from raster to create a list and merge elements together into a dataframe
```{r}
#extract rgb pixel values within training polygons and merge them into a dataframe
rgb_values <- exactextractr::exact_extract(site6, site6_poly, include_cols = "className")
rgb_values_tibble <- as_tibble(rgb_values[[1]])

for (i in 3:length(site6_poly)) {
  temp <- as_tibble(site6_poly[[i]])
  rgb_values_tibble <- bind_rows(rgb_values, temp)
}
#only include pixels that are at least 95% within a polygon
rgb_values_tibble_final <- rgb_values_tibble %>%
  filter(coverage_fraction >= 0.95)
#remove all NA values from the tibble
rgb_values_tibble_final_nonas <- rgb_values_tibble_final %>%
  tidyr::drop_na()
```

```{r}
#convert the site6 raster to an array
site6_array <- as.array(site6)
#calculate the number of edges 
site6_gray <- rgb_2gray(site6_array)
site6_gray[is.na(site6_gray)] <- mean(site6_gray, na.rm = TRUE)
edges <- edge_detection(site6_gray, method = "Sobel")
num_edges <- sum(edges)
```

```{r}
#add a new column for the number of edges
rgb_values_tibble_final_nonas$edges <- num_edges
#remove the geometry and coverage_fraction column as we dont need them for the learner
rgb_values_tibble_final_nonas <- select(rgb_values_tibble_final_nonas, -geometry)
rgb_values_tibble_final_nonas <- select(rgb_values_tibble_final_nonas, -coverage_fraction)
#rename the tibble to "data" 
data <- rgb_values_tibble_final_nonas
```

```{r}
#convert the className column to a factor so it can be used in the mlr3 task. 
data$className <- as.factor(data$className)
data <- na.omit(data)
```

#mlr3
Create mlr3 task
```{r}
task <- TaskClassif$new(id = "test_classification", backend = data, target = "className")
```

Create random forest learner
```{r}
#create a random forest learner
learner <- lrn("classif.ranger", predict_type = "response")
```

Split data into 70% training and 30% validation
```{r}
set.seed(123)
train <- sample(task$nrow, 0.7 * task$nrow)
test <- setdiff(seq_len(task$nrow), train)
```

Train the random forest model
```{r}
learner$train(task, row_ids = train)
```

Make predictions on the test data
```{r}
pred <- learner$predict(task, row_ids = test)
```
Evaluate model and obtain accuracy and kappa value
```{r}
conmat <- pred$confusion

#calculate overall accuracy
n <- sum(conmat)
diag <- diag(conmat)
OA <- sum(diag) / n
OA
#calculate kappa value
rowsums <- apply(conmat, 1, sum)
p <- rowsums / n
colsums <- apply(conmat, 2, sum)
q <- colsums / n
expAccuracy <- sum(p*q)
kappa <- (OA - expAccuracy) / (1 - expAccuracy)
kappa
```
##Apply the trained model to the entire raster

Extract rgb values of the raster and add a "className" column to it
```{r}
site6_boundary <- st_read(here("Data", "TestClassification", "site6_cropBoundary_smaller.shp"))
rgb_values_wholesite <- extract(site6, site6_boundary, ID=F, xy=T)
print(rgb_values_wholesite)

#convert output to a dataframe
rgb_values_wholesite_df <- as.data.frame(rgb_values_wholesite)
#add a "className" column
rgb_values_wholesite_df$className <- NA

#add the edges column that we calculated before
rgb_values_wholesite_df$edges <- num_edges
print(rgb_values_wholesite_df)
```

```{r}
#remove NA values in "red" column
rgb_values_wholesite_df <- rgb_values_wholesite_df[!is.na(rgb_values_wholesite_df$red), ]
#convert "className" column into a factor
rgb_values_wholesite_df$className <- as.factor(rgb_values_wholesite_df$className)
print(rgb_values_wholesite_df)
```

Apply the model to the new dataframe
```{r}
predict <- learner$predict_newdata(newdata = rgb_values_wholesite_df)
```


