---
title: "classification_test_5"
author: "Natalie"
date: "2024-07-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load packages}
library(mlr3)
library(mlr3learners)
library(mlr3spatial)
library(terra)
library(sf)
library(ggplot2)
library(tidyterra)
library(here)
library(mlr3cluster)
```

Load in data. 
```{r}
site6_rast <- rast(here("IURA-2024", "Data", "site6_cropped", "site6_smallER.tif"))
names(site6_rast) <- c("Band_1", "Band_2", "Band_3")
site6_rast

#The geopackage contains 200 random points per polygon and spectral information (rgb values) from the raster. 
site6_vect <- read_sf(here("IURA-2024", "Data", "site6_cropped", "site6_cropped_pts_rgb.gpkg"), stringsAsFactors = TRUE)
site6_vect
```
Create a spatial mlr3 task
```{r}
task <- as_task_classif_st(site6_vect, target = "className")
```

Partition the task into 70% train and 30% test sets
```{r}
set.seed(100)
splits <- partition(task, ratio = 0.7)
```

Create and train a random forest learner
```{r}
learner_rf <- mlr3::lrn("classif.ranger")
training_rf <- learner_rf$train(task, splits$train)
```

Test the random forest learner and obtain overall accuracy
```{r}
testing_rf <- learner_rf$predict(task, splits$test)
(testing_rf$score(msr("classif.acc"))
```

Apply the trained model to the entire raster
```{r}
classified <- predict_spatial(site6_rast, learner_rf)
```
```{r}
print(classified)
writeRaster(classified, here::here("IURA-2024", "Data", "site6_cropped", "site6_small_classified.tif"))
```

Plot the classified raster
```{r}
ggplot() +
  geom_spatraster(data = classified) +
  scale_fill_viridis_d(name = "Class", labels = c("Antelope Brush", "Ground", "Vegetation")) +
  theme_minimal()
```
Calculate area of each class 
```{r}
activeCat(classified) <- "categories"
classified
f <- freq(classified)
f$area <- f$count * (0.013 * 0.013) #in m2
f$percent <- round(100 * f$area / sum(f$area), 1)
f
```

# Run same code but on a 5cm raster to compare accuracy

Import raster and random points
```{r}
#Import raster
site6_5cm <- rast(here("Data", "site6_5cm.tif"))
names(site6_5cm) <- c("Band_1", "Band_2", "Band_3")

#iImport points
site6_5cm_vect <- read_sf(here("Data", "site6_5cm_pts_200.gpkg"), stringsAsFactors = TRUE)
site6_5cm_vect
```
Create spatial task
```{r}
task_5cm <- as_task_classif_st(site6_5cm_vect, target = "className")
```

Partition the task into 70% train and 30% test sets
```{r}
set.seed(100)
splits_5cm <- partition(task_5cm, ratio = 0.7)
```

Create and train a random forest learner
```{r}
learner_5cm_rf <- mlr3::lrn("classif.ranger")
training_5cm_rf <- learner_5cm_rf$train(task_5cm, splits_5cm$train)
```

Test the random forest learner and obtain overall accuracy
```{r}
testing_5cm_rf <- learner_5cm_rf$predict(task_5cm, splits_5cm$test)
testing_5cm_rf$score(msr("classif.acc"))
```
Apply the trained model to the entire raster
```{r}
classified_5cm <- predict_spatial(site6_5cm, learner_5cm_rf)
```
Plot the classified raster
```{r}
ggplot() +
  geom_spatraster(data = classified_5cm) +
  scale_fill_viridis_d(name = "Class", labels = c("Antelope Brush", "Ground", "Vegetation")) +
  theme_minimal()
```
Calculate area of each class 
```{r}
activeCat(classified_5cm) <- "categories"
classified_5cm
freq_5cm <- freq(classified_5cm)
freq_5cm$area <- freq_5cm$count * (0.013 * 0.013) #0.013m * 0.013m
freq_5cm$percent <- round(100 * freq_5cm$area / sum(freq_5cm$area), 1)
freq_5cm
```