---
title: "Sites_7cm"
author: "Natalie"
date: "2024-07-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load packages}
library(terra)
library(here)
```

Import site rasters
```{r}
site1 <- terra::rast(here("Data", "Sites_1.3cm", "site1_1.3cm.tif"))
site2 <- terra::rast(here("Data", "Sites_1.3cm", "site2_1.3cm.tif"))
site3 <- terra::rast(here("Data", "Sites_1.3cm", "site3_1.3cm.tif"))
site4 <- terra::rast(here("Data", "Sites_1.3cm", "site4_1.3cm.tif"))
site5 <- terra::rast(here("Data", "Sites_1.3cm", "site5_1.3cm.tif"))
site6 <- terra::rast(here("Data", "Sites_1.3cm", "site6_1.3cm.tif"))
site7 <- terra::rast(here("Data", "Sites_1.3cm", "site7_1.3cm.tif"))
site8 <- terra::rast(here("Data", "Sites_1.3cm", "site8_1.3cm.tif"))
```

# Changing raster resolution to 7cm/pixel
Rasters are all currently 1.3cm/pixel and we will resample them using bilinear interpolation. Each site will have the same workflow of creating a template with the site's extent and the new resolution, resampling the site raster and then saving the resampled site raster as a .tif file. 

Set new resolution of 7cm/pixel
```{r}
new_res <- 0.07 #7cm = 0.07m
```

Site 1
```{r}
template1 <- rast(ext(site1), res = new_res)
site1_7cm <- resample(site1, template1, method = "bilinear")
writeRaster(site1_7cm, here("Data", "Sites_7cm", "site1_7cm.tif"), overwrite = TRUE)
```
Site 2
```{r}
template2 <- rast(ext(site2), res = new_res)
site2_7cm <- resample(site2, template2, method = "bilinear")
writeRaster(site2_7cm, here("Data", "Sites_7cm", "site2_7cm.tif"), overwrite = TRUE)
```
Site 3
```{r}
template3 <- rast(ext(site3), res = new_res)
site3_7cm <- resample(site3, template3, method = "bilinear")
writeRaster(site3_7cm, here("Data", "Sites_7cm", "site3_7cm.tif"), overwrite = TRUE)
```
Site 4
```{r}
template4 <- rast(ext(site4), res = new_res)
site4_7cm <- resample(site4, template4, method = "bilinear")
writeRaster(site4_7cm, here("Data", "Sites_7cm", "site4_7cm.tif"), overwrite = TRUE)
```
Site 5
```{r}
template5 <- rast(ext(site5), res = new_res)
site5_7cm <- resample(site5, template5, method = "bilinear")
writeRaster(site5_7cm, here("Data", "Sites_7cm", "site5_7cm.tif"), overwrite = TRUE)
```
Site 6
```{r}
template6 <- rast(ext(site6), res = new_res)
site6_7cm <- resample(site6, template6, method = "bilinear")
writeRaster(site6_7cm, here("Data", "Sites_7cm", "site6_7cm.tif"), overwrite = TRUE)
```
Site 7
```{r}
template7 <- rast(ext(site7), res = new_res)
site7_7cm <- resample(site7, template7, method = "bilinear")
writeRaster(site7_7cm, here("Data", "Sites_7cm", "site7_7cm.tif"), overwrite = TRUE)
```
Site 8
```{r}
template8 <- rast(ext(site8), res = new_res)
site8_7cm <- resample(site8, template8, method = "bilinear")
writeRaster(site8_7cm, here("Data", "Sites_7cm", "site8_7cm.tif"), overwrite = TRUE)
```

