---
title: "site2_rgb_edges"
author: "Natalie"
date: "2024-07-31"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Load packages, message=F}
library(mlr3)
library(mlr3learners)
library(mlr3spatial)
library(terra)
library(sf)
library(ggplot2)
library(tidyterra)
library(here)
library(mlr3cluster)
library(OpenImageR)
library(EBImage)
```

Import Datasets
```{r}
site2_rast <- rast(here("Data", "site2_cropped", "site2_cropped.tif"))
names(site2_rast) <- c("Band_1", "Band_2", "Band_3")
plotRGB(site2_rast)

site2_pts <- read_sf(here("Data", "site2_cropped", "site2_cropped_pts.shp"))
site2_pts <- select(site2_pts, -rand_point)
site2_pts
```

Convert the raster to an array
```{r}
site2_array <- as.array(site2_rast)
```

Calculate edges per pixel
```{r}
site2_gray <- rgb_2gray(site2_array)
site2_gray[is.na(site2_gray)] <- mean(site2_gray, na.rm = TRUE)
site2_edges <- edge_detection(site2_gray, method = "Sobel")
```

Convert edges to a raster
```{r}
site2_edges_rast <- rast(site2_edges)
ext(site2_edges_rast) <- ext(site2_rast)
crs(site2_edges_rast) <- crs(site2_rast)
names(site2_edges_rast) <- "edges"
plot(site2_edges_rast)
```
Add edges layer to existing site raster
```{r}
site2_rgb_edges_rast <- c(site2_rast, site2_edges_rast)
site2_rgb_edges_rast
```
Extract raster values
```{r}
site2_rgb_edges_vals <- extract(site2_rgb_edges_rast, site2_pts)
site2_rgb_edges_vals
site2_rgb_edges_vect <- cbind(site2_pts, site2_rgb_edges_vals)
site2_rgb_edges_vect
```
Convert vector to a geopackage
```{r}
site2_rgb_edges_geopkg <- here("Data", "site2_cropped", "site2_rgb_edges_geopkg.gpkg")
st_write(site2_rgb_edges_vect, site2_rgb_edges_geopkg, filetype="GPKG", append=F)
site2_rgb_edges_geopkg <- read_sf(here("Data", "site2_cropped", "site2_rgb_edges_geopkg.gpkg"), stringsAsFactors = T)
site2_rgb_edges_geopkg <- select(site2_rgb_edges_geopkg, -ID)
site2_rgb_edges_geopkg
```
Create spatial task using mlr3
```{r}
site2_rgb_edges_tsk <- as_task_classif_st(site2_rgb_edges_geopkg, target = "className")
```

Partition data into 70% training and 30% testing
```{r}
set.seed(102)
site2_rgb_edges_split <- partition(site2_rgb_edges_tsk, ratio = 0.7)
```

Create random forest learner
```{r}
site2_rgb_edges_lrn <- lrn("classif.ranger")
```

Train learner using the "training" partition
```{r}
site2_rgb_edges_train <- site2_rgb_edges_lrn$train(site2_rgb_edges_tsk, site2_rgb_edges_split$train)
```

Use the "test" partition to obtain overall accuracy of the model
```{r}
site2_rgb_edges_test <- site2_rgb_edges_lrn$predict(site2_rgb_edges_tsk, site2_rgb_edges_split$test)
site2_rgb_edges_test$score(msr("classif.acc"))
```

Apply the trained learner to the entire raster
```{r}
site2_rgb_edges_classified <- predict_spatial(site2_rgb_edges_rast, site2_rgb_edges_lrn)
```

Plot the classified raster
```{r}
ggplot() +
  geom_spatraster(data = site2_rgb_edges_classified) +
  scale_fill_viridis_d(name = "Class", labels = c("Antelope Brush", "Ground", "Vegetation")) +
  theme_minimal()
```
Calculate the area of each class
```{r}
activeCat(site2_rgb_edges_classified) <- "categories"
f <- freq(site2_rgb_edges_classified)
f$area <- f$count * (0.013 * 0.013) #in m2
f$percent <- round(100 * f$area / sum(f$area), 1)
f
```
Export classified raster
```{r}
writeRaster(site2_rgb_edges_classified, here("Data", "site2_cropped", "site2_cropped_classified.tif"), filetype = "GTiff")
```

