---
title: "site4_rgb_edge_test"
author: "Natalie"
date: "2024-07-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Load packages, message=F}
library(mlr3)
library(mlr3learners)
library(mlr3spatial)
library(terra)
library(sf)
library(ggplot2)
library(tidyterra)
library(here)
library(mlr3cluster)
library(OpenImageR)
library(EBImage)
```

Import Datasets
```{r}
site4_rast <- rast(here("Data", "site4_cropped", "site4_cropped.tif"))
names(site4_rast) <- c("Band_1", "Band_2", "Band_3")
plotRGB(site4_rast)

site4_pts <- read_sf(here("Data", "site4_cropped", "site4_randomPts.gpkg"), stringsAsFactors = T)
keep <- c("className", "geom")
site4_pts <- site4_pts[, keep]
site4_pts
```
Convert the raster to an array
```{r}
site4_array <- as.array(site4_rast)
```

Calculate edges per pixel
```{r}
site4_gray <- rgb_2gray(site4_array)
site4_gray[is.na(site4_gray)] <- mean(site4_gray, na.rm = TRUE)
site4_edges <- edge_detection(site4_gray, method = "Sobel")
```

Convert edges to a raster
```{r}
site4_edges_rast <- rast(site4_edges)
ext(site4_edges_rast) <- ext(site4_rast)
crs(site4_edges_rast) <- crs(site4_rast)
names(site4_edges_rast) <- "edges"
plot(site4_edges_rast)
```
Add edges layer to existing site raster
```{r}
site4_rgb_edges_rast <- c(site4_rast, site4_edges_rast)
site4_rgb_edges_rast
```
Extract raster values at random points
```{r}
site4_rgb_edges_vect <- vect(site4_pts)
site4_rgb_edges_vals <- extract(site4_rgb_edges_rast, site4_rgb_edges_vect)
site4_rgb_edges_vect <- cbind(site4_rgb_edges_vect, site4_rgb_edges_vals)
```

Convert SpatVector back to Geopackage
```{r}
site4_rgb_edges_geopkg <- here("Data", "site4_cropped", "site4_rgb_edges_geopkg.gpkg")
writeVector(site4_rgb_edges_vect, site4_rgb_edges_geopkg, filetype="GPKG")
site4_rgb_edges_geopkg <- read_sf(here("Data", "site4_cropped", "site4_rgb_edges_geopkg.gpkg"), stringsAsFactors = T)
site4_rgb_edges_geopkg
```
Remove the ID column as its un-needed
```{r}
site4_rgb_edges_geopkg <- select(site4_rgb_edges_geopkg, -ID)
site4_rgb_edges_geopkg
```
Create spatial task using mlr3
```{r}
site4_rgb_edges_tsk <- as_task_classif_st(site4_rgb_edges_geopkg, target = "className")
```

Partition data into 70% training and 30% testing
```{r}
set.seed(101)
site4_rgb_edges_split <- partition(site4_rgb_edges_tsk, ratio = 0.7)
```

Create random forest learner
```{r}
site4_rgb_edges_lrn <- lrn("classif.ranger")
```

Train learner using the "training" partition
```{r}
site4_rgb_edges_train <- site4_rgb_edges_lrn$train(site4_rgb_edges_tsk, site4_rgb_edges_split$train)
```

Use the "test" partition to obtain overall accuracy of the model
```{r}
site4_rgb_edges_test <- site4_rgb_edges_lrn$predict(site4_rgb_edges_tsk, site4_rgb_edges_split$test)
site4_rgb_edges_test$score(msr("classif.acc"))
```

Apply the trained learner to the entire raster
```{r}
site4_rgb_edges_classified <- predict_spatial(site4_rgb_edges_rast, site4_rgb_edges_lrn)
```

Plot the classified raster
```{r}
ggplot() +
  geom_spatraster(data = site4_rgb_edges_classified) +
  scale_fill_viridis_d(name = "Class", labels = c("Antelope Brush", "Ground", "Vegetation")) +
  theme_minimal()
```
Export classified raster
```{r}
writeRaster(site4_rgb_edges_classified, here("Data", "site2_cropped", "site2_cropped_classified.tif"), filetype = "GTiff")
```
