---
title: "Sites_LowerRes"
author: "Natalie"
date: "2024-07-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load packages}
library(terra)
library(here)
```

Import site rasters
```{r}
site1 <- terra::rast(here("Data", "Sites", "site1_new.tif"))
site2 <- terra::rast(here("Data", "Sites", "site2_new.tif"))
site3 <- terra::rast(here("Data", "Sites", "site3_new.tif"))
site4 <- terra::rast(here("Data", "Sites", "site4_new.tif"))
site5 <- terra::rast(here("Data", "Sites", "site5_new.tif"))
site6 <- terra::rast(here("Data", "Sites", "site6_new.tif"))
site7 <- terra::rast(here("Data", "Sites", "site7_new.tif"))
site8 <- terra::rast(here("Data", "Sites", "site8_new.tif"))
```

# Changing raster resolution to 5cm/pixel
Rasters are all currently 1.3cm/pixel and we will resample them using bilinear interpolation. Each site will have the same workflow of creating a template with the site's extent and the new resolution, resampling the site raster and then saving the resampled site raster as a .tif file. 

Set new resolution of 5cm/pixel
```{r}
new_res <- 0.05 #5cm = 0.05m
```

Site 1
```{r}
template1 <- rast(ext(site1), res = new_res)
site1_5cm <- resample(site1, template1, method = "bilinear")
writeRaster(site1_5cm, here("Data", "Sites_5cm", "site1_5cm.tif"), overwrite = TRUE)
```
Site 2
```{r}
template2 <- rast(ext(site2), res = new_res)
site2_5cm <- resample(site2, template2, method = "bilinear")
writeRaster(site2_5cm, here("Data", "Sites_5cm", "site2_5cm.tif"), overwrite = TRUE)
```
Site 3
```{r}
template3 <- rast(ext(site3), res = new_res)
site3_5cm <- resample(site3, template3, method = "bilinear")
writeRaster(site3_5cm, here("Data", "Sites_5cm", "site3_5cm.tif"), overwrite = TRUE)
```
Site 4
```{r}
template4 <- rast(ext(site4), res = new_res)
site4_5cm <- resample(site4, template4, method = "bilinear")
writeRaster(site4_5cm, here("Data", "Sites_5cm", "site4_5cm.tif"), overwrite = TRUE)
```
Site 5
```{r}
template5 <- rast(ext(site5), res = new_res)
site5_5cm <- resample(site5, template5, method = "bilinear")
writeRaster(site5_5cm, here("Data", "Sites_5cm", "site5_5cm.tif"), overwrite = TRUE)
```
Site 6
```{r}
template6 <- rast(ext(site6), res = new_res)
site6_5cm <- resample(site6, template6, method = "bilinear")
writeRaster(site6_6cm, here("Data", "Sites_5cm", "site6_5cm.tif"), overwrite = TRUE)
```
Site 7
```{r}
template7 <- rast(ext(site7), res = new_res)
site7_5cm <- resample(site7, template7, method = "bilinear")
writeRaster(site7_5cm, here("Data", "Sites_5cm", "site7_5cm.tif"), overwrite = TRUE)
```
Site 8
```{r}
template8 <- rast(ext(site8), res = new_res)
site8_5cm <- resample(site8, template8, method = "bilinear")
writeRaster(site8_5cm, here("Data", "Sites_5cm", "site8_5cm.tif"), overwrite = TRUE)
```

