---
title: "site4_rgb_test"
author: "Natalie"
date: "2024-07-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Load packages, message=F}
library(mlr3)
library(mlr3learners)
library(mlr3spatial)
library(terra)
library(sf)
library(ggplot2)
library(tidyterra)
library(here)
library(mlr3cluster)
library(OpenImageR)
library(EBImage)
```

Import datasets
```{r}
site4_rast <- rast(here("Data", "site4_cropped", "site4_cropped.tif"))
names(site4_rast) <- c("Band_1", "Band_2", "Band_3")
plotRGB(site4_rast)

site4_pts <- read_sf(here("Data", "site4_cropped", "site4_randomPts.gpkg"), stringsAsFactors = T)
keep <- c("className", "geom")
site4_pts <- site4_pts[, keep]
site4_pts
```

Extract raster values at the random points
```{r}
#convert gpkg to spatvector
site4_vect <- vect(site4_pts)
#extract rgb vals
site4_rgb_vals <- extract(site4_rast, site4_vect)
site4_rgb <- cbind(site4_vect, site4_rgb_vals)
#convert the spatvector back into a geopackage containing its geometry
site4_rgb_geopkg <- here("Data", "site4_cropped", "site4_rgb_geopkg.gpkg")
writeVector(site4_rgb, site4_rgb_geopkg, filetype="GPKG")
site4_rgb_geopkg <- read_sf(here("Data", "site4_cropped", "site4_rgb_geopkg.gpkg"), stringsAsFactors = T)
site4_rgb_geopkg
```
Remove the ID column as its un-needed
```{r}
site4_rgb_geopkg <- select(site4_rgb_geopkg, -ID)
site4_rgb_geopkg
```
Create a spatial task using mlr3
```{r}
site4_rgb_tsk <- as_task_classif_st(site4_rgb_geopkg, target="className") 
```

Partition data into 70% training and 30% testing
```{r}
set.seed(101)
site4_split <- partition(site4_rgb_tsk, ratio = 0.7)
```

Create random forest learner
```{r}
site4_rgb_lrn <- lrn("classif.ranger")
```

Train learner using the "training" partition
```{r}
site4_rgb_train <- site4_rgb_lrn$train(site4_rgb_tsk, site4_split$train)
```

Use the "test" partition to obtain overall accuracy of the model
```{r}
site4_rgb_test <- site4_rgb_lrn$predict(site4_rgb_tsk, site4_split$test)
site4_rgb_test$score(msr("classif.acc"))
```

Apply the trained learner to the entire raster
```{r}
site4_rgb_classified <- predict_spatial(site4_rast, site4_rgb_lrn)
```

Plot the classified raster
```{r}
ggplot() +
  geom_spatraster(data = site4_rgb_classified) +
  scale_fill_viridis_d(name = "Class", labels = c("Antelope Brush", "Ground", "Vegetation")) +
  theme_minimal()
```
Export the classified raster 
```{r}
writeRaster(site4_rgb_classified, here("Data", "site4_cropped", "site4_cropped_classified.tif"), filetype = "GTiff")
```

Calculate the area of each class
```{r}
activeCat(site4_rgb_classified) <- "categories"
f <- freq(site4_rgb_classified)
f$area <- f$count * (0.013 * 0.013) #in m2
f$percent <- round(100 * f$area / sum(f$area), 1)
f
```

